---
title: "BBC Style Graphics"
author: "Salerno"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(bbplot)
```

## 1. Libraries that you need

```{r warning=FALSE, echo=FALSE}

#This line of code installs the pacman page if you do not have it installed - if you do, it simply loads the package
if(!require(pacman))install.packages("pacman")

pacman::p_load('bbplot',
              'dplyr',
              'forcats',
              'gapminder',
              'ggalt',
              'ggplot2',
              'ggpubr',
              'grid',
              'png',
              'R.utils',
              'scales',
              'tidyr') 

```

## 2. bbplot

```{r warning=TRUE, echo=FALSE}

# install.packages('devtools')
devtools::install_github('bbc/bbplot', force = TRUE)

```

## 3. gapminder example using BBC template

```{r}

#Data for chart from gapminder package
line_df <- gapminder %>%
  filter(country == "Brazil") 

#Make plot
line <- ggplot(line_df, aes(x = year, y = lifeExp)) +
  geom_line(colour = "#1380A1", size = 1) +
  geom_hline(yintercept = 0, size = 1, colour="#333333") +
  bbc_style() +
  labs(title="Living longer",
       subtitle = "Life expectancy in Brazil 1952-2007") +
  theme(panel.grid.major.x = element_line(color="#cbcbcb"), 
        panel.grid.major.y=element_blank())

```

## 4. Finalize your plot

```{r warning=TRUE}
library(ggplot2)
library(magick)
library(here) # For making the script run without a wd
library(magrittr) # For piping the logo

# Make a simple plot and save it

living_plot <- finalise_plot(plot_name = line,
                             source = "Source: Gapminder",
                             save_filepath = "BBC_gapminder.png",
                             width_pixels = 640,
                             height_pixels = 450)

# Call back the plot
plot <- image_read(paste0(getwd(), "/BBC_gapminder.png"))


# Scale down the logo and give it a border and annotation
# This is the cool part because you can do a lot to the image/logo before adding it
logo <- function(){
  # And bring in a logo
  logo_raw <- image_read(paste0(getwd(), "/data_fintech_solutions_logo.png"))
  
  logo_raw %>%
  image_scale("500") %>% 
  image_background("grey", flatten = TRUE) %>%
  image_border("grey", "400x1") %>%
  image_annotate("Powered By R", color = "white", size = 35, 
                 location = "+10+50", gravity = "southeast")
  
}


# Stack them on top of each other
final_plot <- image_append(image_scale(c(plot, logo()), "500"), stack = TRUE)
# And overwrite the plot without a logo
image_write(final_plot, paste0(getwd(), last_plot()$labels$title, ".png"))

```

## 5. Make a line chart

```{r}

#Prepare data
line_df <- gapminder %>%
  filter(country == "China") 

#Make plot
line <- ggplot(line_df, aes(x = year, y = lifeExp)) +
  geom_line(colour = "#1380A1", size = 1) +
  geom_hline(yintercept = 0, size = 1, colour="#333333") +
  bbc_style() +
  labs(title="Living longer",
       subtitle = "Life expectancy in China 1952-2007")

line
```

## 6. Making a multiple line chart

```{r}

#Prepare data
multiple_line_df <- gapminder %>%
  filter(country == "China" | country == "United States") 

#Make plot
multiple_line <- ggplot(multiple_line_df, aes(x = year, y = lifeExp, colour = country)) +
  geom_line(size = 1) +
  geom_hline(yintercept = 0, size = 1, colour="#333333") +
  scale_colour_manual(values = c("#FAAB18", "#1380A1")) +
  bbc_style() +
  labs(title="Living longer",
       subtitle = "Life expectancy in China and the US")

multiple_line
```

## 7. Making a bar chart

```{r}

#Prepare data
bar_df <- gapminder %>%
  filter(year == 2007 & continent == "Africa") %>%
  arrange(desc(lifeExp)) %>%
  head(5)

#Make plot
bars <- ggplot(bar_df, aes(x = country, y = lifeExp)) +
  geom_bar(stat="identity", 
           position="identity", 
           fill="#1380A1") +
  geom_hline(yintercept = 0, size = 1, colour="#333333") +
  bbc_style() +
  labs(title="Reunion is highest",
       subtitle = "Highest African life expectancy, 2007")

bars
```

## 8. Making a stacked bar chart

```{r}

#prepare data
stacked_df <- gapminder %>% 
  filter(year == 2007) %>%
  mutate(lifeExpGrouped = cut(lifeExp, 
                    breaks = c(0, 50, 65, 80, 90),
                    labels = c("Under 50", "50-65", "65-80", "80+"))) %>%
  group_by(continent, lifeExpGrouped) %>%
  summarise(continentPop = sum(as.numeric(pop)))

#set order of stacks by changing factor levels
stacked_df$lifeExpGrouped = factor(stacked_df$lifeExpGrouped, levels = rev(levels(stacked_df$lifeExpGrouped)))

#create plot
stacked_bars <- ggplot(data = stacked_df, 
                       aes(x = continent,
                           y = continentPop,
                           fill = lifeExpGrouped)) +
  geom_bar(stat = "identity", 
           position = "fill") +
  bbc_style() +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_viridis_d(direction = -1) +
  geom_hline(yintercept = 0, size = 1, colour = "#333333") +
  labs(title = "How life expectancy varies",
       subtitle = "% of population by life expectancy band, 2007") +
  theme(legend.position = "top", 
        legend.justification = "left") +
  guides(fill = guide_legend(reverse = TRUE))

stacked_bars

```

## 9. Making a grouped bar chart

```{r}

#Prepare data
grouped_bar_df <- gapminder %>%
  filter(year == 1967 | year == 2007) %>%
  select(country, year, lifeExp) %>%
  spread(year, lifeExp) %>%
  mutate(gap = `2007` - `1967`) %>%
  arrange(desc(gap)) %>%
  head(5) %>%
  gather(key = year, 
         value = lifeExp,
         -country,
         -gap) 
  
#Make plot
grouped_bars <- ggplot(grouped_bar_df, 
                       aes(x = country, 
                           y = lifeExp, 
                           fill = as.factor(year))) +
  geom_bar(stat="identity", position="dodge") +
  geom_hline(yintercept = 0, size = 1, colour="#333333") +
  bbc_style() +
  scale_fill_manual(values = c("#1380A1", "#FAAB18")) +
  labs(title="We're living longer",
       subtitle = "Biggest life expectancy rise, 1967-2007")

grouped_bars

```

## 10. Making a dumbbell chart

```{r}

library("ggalt")
library("tidyr")

#Prepare data
dumbbell_df <- gapminder %>%
  filter(year == 1967 | year == 2007) %>%
  select(country, year, lifeExp) %>%
  spread(year, lifeExp) %>%
  mutate(gap = `2007` - `1967`) %>%
  arrange(desc(gap)) %>%
  head(10)

#Make plot
ggplot(dumbbell_df, aes(x = `1967`, xend = `2007`, y = reorder(country, gap), group = country)) + 
  geom_dumbbell(colour = "#dddddd",
                size = 3,
                colour_x = "#FAAB18",
                colour_xend = "#1380A1") +
  bbc_style() + 
  labs(title="We're living longer",
       subtitle="Biggest life expectancy rise, 1967-2007")

```

## 11. Making a histogram

```{r}

hist_df <- gapminder %>%
  filter(year == 2007)

ggplot(hist_df, aes(lifeExp)) +
  geom_histogram(binwidth = 5, colour = "white", fill = "#1380A1") +
  geom_hline(yintercept = 0, size = 1, colour="#333333") +
  bbc_style() +
  scale_x_continuous(limits = c(35, 95),
                     breaks = seq(40, 90, by = 10),
                     labels = c("40", "50", "60", "70", "80", "90 years")) +
  labs(title = "How life expectancy varies",
       subtitle = "Distribution of life expectancy in 2007")

```

## 12. Makes changes to the legend

```{r}

multiple_line + guides(colour=FALSE)

```

### 12.1 Removing all legends

```{r}

multiple_line + theme(legend.position = "none")

```

### 12.2 Changing position of the legend

```{r}

multiple_line + theme(legend.position = "right")

```

### 12.3 Definig an exact legend position

```{r}

multiple_line + theme(legend.position = c(0.115,1.05),
                      legend.direction = "horizontal") +  
  labs(title="Living longer",
       subtitle = "Life expectancy in China and the US\n") +
  theme(legend.margin = margin(0, 0, 0, 100))


```

### 12.4 Remove the legend title

```{r}

#Prepare data
multiple_line_df <- gapminder %>%
  filter(country == "China" | country == "United States") 

#Make plot
multiple_line <- ggplot(multiple_line_df, aes(x = year, y = lifeExp, colour = country)) +
  geom_line(size = 1) +
  geom_hline(yintercept = 0, size = 1, colour="#333333") +
  scale_colour_manual(values = c("#FAAB18", "#1380A1")) +
  bbc_style() +
  theme(legend.title = element_blank()) +
  labs(title="Living longer",
       subtitle = "Life expectancy in China and the US")
  

multiple_line

```

### 12.5 Reversing the order

```{r}

multiple_line + guides(fill = guide_legend(reverse = TRUE))

```

### 12.6 Rearranging the layout of your legend

```{r}

multiple_line + guides(fill = guide_legend(nrow = 4, byrow = T))

```

### 12.7 Changing the apperarance of your leged symbols

```{r}

multiple_line + guides(fill = guide_legend(override.aes = list(size = 4)))

```

### 12.8 Adding space between your legends labels

```{r}

multiple_line + scale_colour_manual(labels = function(x) paste0(" ", x, "  "))

```

## 13. Making changes to the axes

```{r}

bars <- bars + coord_flip()

bars

```

### 13.1 Adding/removing gridlines

```{r}

bars <- bars + coord_flip() +
  theme(panel.grid.major.x = element_line(color="#cbcbcb"), 
        panel.grid.major.y=element_blank())

bars
```

### 13.2 Changing the axis text manually

```{r}
bars <- bars + scale_y_continuous(limits=c(0,85),
                   breaks = seq(0, 80, by = 20),
                   labels = c("0","20", "40", "60", "80 years"))

bars


```

### 13.3 Addinh thousand separators to your axis labels

```{r}

bars + ggplot2::scale_y_continuous(labels = function(x) format(x, big.mark = ",",
                                                 scientific = FALSE))

```

### 13.4 Adding percent symbol to your axis labels

```{r}

bars + ggplot2:::scale_y_continuous(labels = function(x) paste0(x, "%"))

```

### 13.5 Changing the plot limits

```{r}

bars + ggplot2::ylim(c(0,500))

## Scale for 'y' is already present. Adding another scale for 'y', which
## will replace the existing scale.

```

### 13.6 Adding axis titles

```{r}

bars + ggplot2::theme(axis.title = ggplot2::element_text(size = 18))


```

### 13.7 Modifying axis title

```{r}

bars + ggplot2::labs(x = "I'm an axis", 
       y = "")

```

### 13.8 Adding axis ticks

```{r}

multiple_line + ggplot2::theme(
  axis.ticks.x = ggplot2::element_line(colour = "#333333"), 
  axis.ticks.length =  ggplot2::unit(0.26, "cm"))


```

### 13.9 Adding annotations

```{r}

multiple_line + ggplot2::geom_label(ggplot2::aes(x = 1980, y = 45, label = "I'm an annotation!"), 
                           hjust = 0, 
                           vjust = 0.5, 
                           colour = "#555555", 
                           fill = "white", 
                           label.size = NA, 
                           family="Helvetica", 
                           size = 6)

```

```{r}

multiple_line <- multiple_line + 
  ggplot2::geom_label(ggplot2::aes(x = 1980, y = 45, label = "I'm quite a long\nannotation over\nthree rows"), 
             hjust = 0, 
             vjust = 0.5, 
             lineheight = 0.8,
             colour = "#555555", 
             fill = "white", 
             label.size = NA, 
             family="Helvetica", 
             size = 6) 

multiple_line
```

```{r}

library(ggplot2)
multiple_line <- multiple_line + 
  theme(legend.position = "none") + 
  xlim(c(1950, 2011)) +
  geom_label(aes(x = 2007, y = 79, label = "US"), 
             hjust = 0, 
             vjust = 0.5, 
             colour = "#1380A1", 
             fill = "white", 
             label.size = NA, 
             family="Helvetica", 
             size = 6) +
  geom_label(aes(x = 2007, y = 72, label = "China"), 
             hjust = 0, 
             vjust = 0.5, 
             colour = "#FAAB18", 
             fill = "white", 
             label.size = NA, 
             family="Helvetica", 
             size = 6)

```

## 14. Left-align/right-align text

```{r}

labelled.bars <- bars +
  geom_label(aes(x = country, y = lifeExp, label = round(lifeExp, 0)),
             hjust = 1, 
             vjust = 0.5, 
             colour = "white", 
             fill = NA, 
             label.size = NA, 
             family="Helvetica", 
             size = 6)

labelled.bars

```

### 14.1 Add left-aligned labels to bar charts

```{r}

labelled.bars.v2 <- bars +
  geom_label(aes(x = country, 
                 y = 4, 
                 label = round(lifeExp, 0)),
             hjust = 0, 
             vjust = 0.5, 
             colour = "white", 
             fill = NA, 
             label.size = NA, 
             family="Helvetica", 
             size = 6)

labelled.bars.v2

```

### 14.2 Add a line

```{r}

multiple_line + geom_segment(aes(x = 1979, y = 45, xend = 1965, yend = 43), 
                             colour = "#555555", 
                             size=0.5)

```

### 14.3 Add a curved line

```{r}

multiple_line + geom_curve(aes(x = 1979, y = 45, xend = 1965, yend = 43), 
                           colour = "#555555", 
                           curvature = -0.2,
                           size=0.5)

```

### 14.4 Add a arrow

```{r}

multiple_line + geom_curve(aes(x = 1979, y = 45, xend = 1965, yend = 43), 
                             colour = "#555555", 
                             size=0.5, 
                             curvature = -0.2,
                             arrow = arrow(length = unit(0.03, "npc")))

```

### 14.5 Add a line across the whole plot

```{r}

multiple_line + geom_hline(yintercept = 10, size = 1, colour = "red", linetype = "dashed")

```

## 15 Work with small multiples

```{r}
library(dplyr)
library(gapminder)

#Prepare data
facet <- gapminder %>%
  filter(continent != "Americas") %>%
  group_by(continent, year) %>%
  summarise(pop = sum(as.numeric(pop)))

#Make plot
facet_plot <- ggplot() +
  geom_area(data = facet, aes(x = year, y = pop, fill = continent)) +
  scale_fill_manual(values = c("#FAAB18", "#1380A1","#990000", "#588300")) + 
  facet_wrap( ~ continent, ncol = 5) + 
  scale_y_continuous(breaks = c(0, 2000000000, 4000000000),
                     labels = c(0, "2bn", "4bn")) +
  bbc_style() +
  geom_hline(yintercept = 0, size = 1, colour = "#333333") +
  theme(legend.position = "none",
        axis.text.x = element_blank()) +
  labs(title = "Asia's rapid growth",
       subtitle = "Population growth by continent, 1952-2007")

facet_plot
```

### 15.1 Free scales

```{r}

#Make plot
facet_plot_free <- ggplot() +
  geom_area(data = facet, aes(x = year, y = pop, fill = continent)) +
  facet_wrap(~ continent, scales = "free") + 
  bbc_style() +
  scale_fill_manual(values = c("#FAAB18", "#1380A1","#990000", "#588300")) +
  geom_hline(yintercept = 0, size = 1, colour = "#333333") +
  theme(legend.position = "none",
        axis.text.x = element_blank(),
        axis.text.y = element_blank()) +
  labs(title = "It's all relative",
       subtitle = "Relative population growth by continent,1952-2007")

facet_plot_free
```

## 16 Increase or decrease margins

```{r}

bars + theme(plot.subtitle=element_text(margin=margin(b=75)))

```

### 16.1 Exporting your plot and x-axis margins

```{r}

bar_chart_tall <- bars + theme(axis.text.x = element_text(margin=margin(t = 7, b = 10)))
bar_chart_tall


```

### 16.2 Reorder bars by size

```{r}

bars <- ggplot(bar_df, aes(x = reorder(country, lifeExp), y = lifeExp)) +
  geom_bar(stat="identity", position="identity", fill="#1380A1") +
  geom_hline(yintercept = 0, size = 1, colour="#333333") +
  bbc_style() +
  coord_flip() +
  labs(title="Reunion is highest",
       subtitle = "Highest African life expectancy, 2007") +
  theme(panel.grid.major.x = element_line(color="#cbcbcb"), 
        panel.grid.major.y=element_blank())

bars
```

### 16.3 Reorder bars manually

```{r}

bars$column <- factor(bars$column, levels = c("18-24","25-64","65+"))

bars <- ggplot(bar_df, aes(x = reorder(bars$column), y = lifeExp)) +
  geom_bar(stat="identity", position="identity", fill="#1380A1") +
  geom_hline(yintercept = 0, size = 1, colour="#333333") +
  bbc_style() +
  coord_flip() +
  labs(title="Reunion is highest",
       subtitle = "Highest African life expectancy, 2007") +
  theme(panel.grid.major.x = element_line(color="#cbcbcb"), 
        panel.grid.major.y=element_blank())

bars

```

### 16.4 Colour bars conditionally

```{r}

ggplot(bar_df, 
       aes(x = reorder(country, lifeExp), y = lifeExp)) +
  geom_bar(stat="identity", position="identity", fill=ifelse(bar_df$country == "Mauritius", "#1380A1", "#dddddd")) +
  geom_hline(yintercept = 0, size = 1, colour="#333333") +
  bbc_style() +
  coord_flip() +
  labs(title="Reunion is highest",
       subtitle = "Highest African life expectancy, 2007") +
  theme(panel.grid.major.x = element_line(color="#cbcbcb"), 
        panel.grid.major.y=element_blank()) 

```

